On the error performance of the $A_n$ lattices

### 2012-03-07 paper.tex
Added beginning of reviewer comments.

%\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
\documentclass[journal, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}
%\documentclass[correspondence,9pt]{IEEEtran}

@@

\begin{abstract}
We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.  These formulae
enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
\end{abstract}

@@

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $(n+1)$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. %It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

@@

The Voronoi cell is linked with the problem of \emph{lattice decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.  Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$). One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
 %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder.
A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}. Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.

@@

Let $\calC$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\cbf \in \calC$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian. However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty}
 \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in
\vor(\Lambda) + \cbf )  \nonumber \\
&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
\begin{align}
P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.  The number of terms required will increase as the noise variance gets smaller.  This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.

\newcommand{\calM}{M}
In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
\[
 \calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
\]
These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.

@@

We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

\begin{theorem}  \label{thm:VorAn=QVorZn1}
The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
\[
\vor(A_n) = \Qbf\vor(\ints^{n+1}).
\]
\end{theorem}
\begin{proof}
Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.  The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.  From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
\[
\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
\]
also. It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
\end{proof}

@@

We would like to be able to integrate functions over the Voronoi cell of $A_n$.  Consider a function $f : \reals^{n+1} \mapsto \reals$.  The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.  The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.  So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.  This is not what we intend.  By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.  However, we find the following approach simpler.  Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.  Now, the integral over the Voronoi cell can be written as
\begin{equation}\label{eq:vorfint}
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
\end{equation}
It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
\begin{equation}\label{eq:unormalised}
\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
\end{equation}
This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.  To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.  Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.  The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
\[
\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
\]
The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
\[
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
\]
The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.

@@

We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables,
\[
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
\]
and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
\begin{equation}\label{eq:intwewilluse}
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\end{equation}
Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
\[
\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
\]
Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.  Interestingly $\wbf$ does not contain $y$ so the term inside the integral~\eqref{eq:intwewilluse} does not depend on $y$.  This is the integral we will use to compute the moments of $A_n$.

\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
\begin{align*}
M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
 &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy. \\
&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1}.
\end{align*}
as required.  Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
\end{example}

>>>>

%\documentclass[a4paper,10pt]{article}
\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}
%\documentclass[correspondence,9pt]{IEEEtran}

@@

\begin{abstract}
We consider the root lattice $A_n$ and derive explicit recursive formulae for the moments of its Voronoi cell.  These formulae
enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
\end{abstract}

@@

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of a $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called a \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as a generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. %It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

@@

The Voronoi cell is linked with the problem of \emph{lattice decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.  Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001}. Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$). One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002,Jalden2005_sphere_decoding_complexity}.
 %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder.
A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}. Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.

@@

Let $\calC$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\cbf \in \calC$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian. However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
%P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty} \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
P_C &\geq P_C(\Lambda) = \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
\begin{align}
P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.  The number of terms required will increase as the noise variance gets smaller.  This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.

\newcommand{\calM}{M}
In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the \emph{moment}s
\[
 \calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
\]
These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.

@@

We now state our main result.  The moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

\begin{theorem}  \label{thm:VorAn=QVorZn1}
The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
\[
\vor(A_n) = \Qbf\vor(\ints^{n+1}).
\]
\end{theorem}
\begin{proof}
Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.  The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.  From Burger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
\[
\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
\]
also. It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
\end{proof}

@@

We would like to be able to integrate functions over the Voronoi cell of $A_n$.  Consider a function $f : \reals^{n+1} \mapsto \reals$.  The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.  The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.  So, the volume integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.  This is not what we intend.  By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.  However, we find the following approach simpler.  Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.  If $S$ has $n$-volume $V$ then the $(n+1)$-volume of $S\times\onebf$ is $V \sqrt{n+1}$.  Now, the integral over the Voronoi cell can be written as
\begin{equation}\label{eq:vorfint}
\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \frac{1}{\sqrt{n+1}} \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
\end{equation}
It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
\begin{equation}\label{eq:unormalised}
\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
\end{equation}
This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.  To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.  Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.  The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of the set of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is
\[
\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
\]
The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
\begin{equation}\label{eq:vorfintafternorm}
\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
\end{equation}
The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.

@@

We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables, gives
\[
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
\]
and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
\begin{equation}\label{eq:intwewilluse}
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\end{equation}
Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
\[
\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
\]
Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.  Interestingly $\wbf$ does not contain $y$ so the term inside the integral~\eqref{eq:intwewilluse} does not depend on $y$.  This is the integral we will use to compute the moments of $A_n$.

\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
\begin{align*}
M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
 &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1},
\end{align*}
as required.  Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
\end{example}
